---
title: "Tidyeval Tutorial"
author: "Methods Consultants of Ann Arbor"
date: "07/13/2017"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    logo: mc_logo_rectangle.png
    css: mc_slides.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)

library(tidyverse)
library(rlang)
library(knitr)

```

## Data

```{r}
data(mtcars)
head(mtcars)
```

## Standard vs Non-Standard Evalulation

Standard evaluation: R passes arguments by value. 

Non-standard evaluation (NSE): Catch-all term that means anything other than standard evaluation.

Dplyr NSE: Allows objects that do not exist in global environment as inputs. 

```{r}
group_by(mtcars, cyl) %>%
  head(n=3)
```

With standard evaluation, cyl cannot be an input as it doesn't exist in global environment. 

## Non-Standard Evalulation Issues

While NSE can be extremely convienient, it causes problems when using non-interactively. 

```{r}
oops <- function(var) {
  mtcars %>%
    group_by(var)
}

oops(cyl)
```

This is where tidyeval comes in.

## Tidyeval

- Tidyeval is a toolkit for the tidyverse

- It was created to help get around the issues that NSE can cause 

- Elegantly functionize dplyr code (other tidyverse packages will have support soon)

## Intro to BangBang

We'll start with a few interactive examples to help get an intuition for how these functions work

```{r}
mtcars %>%
  select(2) %>%
  head(n=3)
```

```{r}
hp <- 2

mtcars %>%
  select(hp) %>%
  head(n=3)
```

## Intro to BangBang

The `!!` operator, pronounced BangBang, tells dplyr to evaluate the object prior to using it. 

```{r}
mtcars %>%
  select(!!hp) %>%
  head(n=3)
```

## Intro to BangBang

`expr()` allows us to debug our tidyeval code, it lets us see what R is trying to evaluate. 

```{r}
expr(
  mtcars %>%
    select(hp) %>%
    head(n=3)
)
```

```{r}
expr(
  mtcars %>%
    select(!!hp) %>%
    head(n=3)
)
```

## Intro to quo and sym

```{r}
mtcars %>%
  filter(gear == 4) %>%
  head(n=3) %>%
  kable()
```

## Intro to quo and sym

```{r}
x <- "gear"

mtcars %>%
  filter(x == 4) %>%
  head(n=3) %>%
  kable()

expr(
  mtcars %>%
    filter(x == 4) %>%
    head(n=3) %>%
    kable()
)
```

Filter ignores that x has an assigned value, "gear". We would like x to be evaluated (replaced with "gear") prior to being used by filter.

## Intro to quo and sym

```{r}
mtcars %>%
  filter((!!x) == 4) %>%
  head(n=3) %>%
  kable()

expr(
  mtcars %>%
    filter((!!x) == 4) %>%
    head(n=3) %>%
    kable()
)
```

`!!` tells dplyr to evaluate `x` a step before evaluating the whole expression

However we don't want a string there, we want the object name to be a symbol. 

## Intro to quo and sym

`sym()` converts string to symbol

```{r}
expr(x)
expr(!!x)
x <- sym(x)
expr(!!x)
```

## Intro to quo and sym

```{r}
x <- sym("gear")

mtcars %>%
  filter((!!x) == 4) %>%
  head(n=3) %>%
  kable()
```

## Intro to quo and sym

Be careful: `filter((!!x) == 4)` is different from `filter(!!x == 4)` 

```{r}
expr(
  mtcars %>%
    filter((!!x) == 4) %>%
    head(n=3) %>%
    kable()
)

expr(
  mtcars %>%
    filter(!!x == 4) %>%
    head(n=3) %>%
    kable()
)
```

## Intro to quo and sym

If we want our input to be unquoted, then we use `quo()`

```{r}
x <- gear
```

R tries to evaluate `gear` and save it into `x`. We know `gear` doesn't exist, and we don't want R to try to evaluate it globally. Just save the expression in `x`!

```{r}
x <- quo(gear)
expr(x)
expr(!!x)
```

## Intro to quo and sym

```{r}
x <- quo(gear)

mtcars %>%
  filter((!!x) == 4) %>%
  head(n=3) %>%
  kable()
```


## Definitions

`!!`: tells R to evaluate the object prior to using it.

`sym()`: takes string input and converts to symbol; since dplyr deprecated the underscore functions, it only takes unquoted inputs (symbols)

quosure: a quosure is a quoted object that keeps track of its environment

`quo()`: takes unquoted R code and returns a quosure ready to be evaluated

`enquo()`: takes symbol referring to function argument, returns quosure of input to function

## tl;dr

***Coding interactively*** (useful while building function): use `sym()` or `quo()`

***Inside function***: use `sym()` or `enquo()`

## Difference between `quo()` and `enquo()`

Say we're trying to group `mtcars` by `cyl`, and need to make a closure of `cyl`

If working interactively:

  - use `quo(cyl)`

If making `function(df, var)`, where user inputs `var = cyl`:

  - use `enquo(var)`
  - `quo(var)` would make closure of `var`

## Enquo and sym example

```{r}
mtcars %>% 
  select(cyl) %>% 
  head(n=3) %>%
  kable()
```

## Enquo and sym example

```{r}

select3 <- function(df, select_var) {

  df %>% 
    select(!!select_var) %>% 
    head(n=3) %>%
    kable()
  
}

mtcars %>% 
  select3(cyl)

```

Tries to evaluate `cyl` globablly, we need to use `enquo()` so that `cyl` can be used correctly by dplyr. 

## Enquo and sym example

```{r}
select3 <- function(df, select_var) {

  select_var <- enquo(select_var)
  
  df %>% 
    select(!!select_var) %>% 
    head(n=3) %>%
    kable()

}

mtcars %>% 
  select3(cyl)
```


## Enquo and sym example

```{r}
select3_string <- function(df, select_var) {
  
  select_var <- sym(select_var)
  
  df %>% 
    select(!!select_var) %>% 
    head(n=3) %>%
    kable()

}

mtcars %>% 
  select3_string("cyl")
```

## Expression instead of variable

```{r}
filter3 <- function(expression) {
  
  expression <- enquo(expression)
  
  mtcars %>%
    filter(!!expression) %>%
    head(n=3) %>%
    kable()

}

filter3(vs == 0)
```

## Dynamic Naming

```{r}
name3 <- function(var) {
  
  var <- enquo(var)
  
  mtcars %>%
    mutate(x = ((!!var) - mean(!!var))/sd(!!var)) %>%
    head(n=3) %>%
    kable()
  
}

name3(mpg)
```

## Dynamic Naming

```{r}
name3 <- function(var) {
  
  var <- enquo(var)
  
  mtcars %>%
    mutate(!!var = ((!!var) - mean(!!var))/sd(!!var)) %>%
    head(n=3) %>%
    kable()
  
}
```

`!!var = ((!!var) - mean(!!var))/sd(!!var)` is not valid R code, the \``=`\` operator can only have object names on the LHS. We need to use \``:=`\` operator to unquote on both RHS and LHS.

## Definitions

\``:=`\`: bundles expressions on both the LHS and RHS. 

`quo_name()`: converts quoted symbol to string

## Dynamic Naming

```{r}
name3 <- function(var) {
  
  var <- enquo(var)
  
  mtcars %>%
    mutate(!!var := ((!!var) - mean(!!var))/sd(!!var)) %>%
    head(n=3) %>%
    kable()
  
}

name3(mpg)
```

## Dynamic Naming

```{r}
name3 <- function(var) {
  
  var <- enquo(var)
  
  name <- quo_name(var)
  
  mtcars %>%
    mutate(!!name := ((!!var) - mean(!!var))/sd(!!var)) %>%
    head(n=3) %>%
    kable()
  
}

name3(mpg)
```

## Dynamic Naming

```{r}
name3 <- function(var) {
  
  var <- enquo(var)
  
  name <- paste0("z_", quo_name(var))
  
  mtcars %>%
    mutate(!!name := ((!!var) - mean(!!var))/sd(!!var)) %>%
    head(n=3) %>%
    kable()
  
}

name3(mpg)
```

## Definitions

`syms()`: takes a list of strings and converts to list of symbols

`quos()`: takes a list of expressions and returns list of quosures

`!!!`: unquotes each element of list, each component is embedded as its own argument in surrounding call

## Quos, syms, !!! example

```{r}

g <- c("gear", "cyl")

multiple_group <- function(df, group_vars) {
  
  group_vars <- syms(group_vars)
  
  df %>%
    group_by(!!!group_vars) %>%
    summarise(mean(hp)) %>%
    head(n=3) %>%
    kable()
  
}

```

## Quos, syms, !!! example

```{r}

multiple_group(mtcars, g)
multiple_group(mtcars, "gear")

```

## Quos, syms, !!! example

```{r}
multiple_group <- function(df, ...) {
  
  val_var <- enquo(val_var)
  groupings <- quos(...)
  
  df %>%
    group_by(!!!groupings) %>%
    summarise(mean(hp)) %>%
    head(n=3) %>%
    kable()
  
}
```

## Quos, syms, !!! example

```{r}
multiple_group(mtcars, gear)
multiple_group(mtcars, gear, cyl)

```

## Quos example 2

```{r}
(example <- tibble(group=c(rep("A",6), rep("B",6)), val = c(1:5, NA, 7:11, 1000)))
```

## Quos example 2

```{r}
grouped_mean <- function(df, ...) {
  
  args <- quos(...)

  df %>%
    group_by(group) %>%
    summarise(value = mean(val, !!!args)) %>%
    kable()
  
}
```

## Quos example 2

```{r}
grouped_mean(example)
grouped_mean(example, na.rm = TRUE)
```

## Quos example 2

```{r}
grouped_mean(example, trim = .2)
grouped_mean(example, na.rm = TRUE, trim = .2)
```

## Allow either string or expression

```{r}
group_either <- function(var) {
  var <- as.character(substitute(var))
  var <- sym(var)
  
  mtcars %>%
    group_by(!!var) %>%
    summarise(mean(hp))
}
```

## Allow either string or expression

```{r}
group_either(cyl)
group_either("cyl")
```

## Final notes

Need to explicitly load rlang if you want to use `expr()`, `sym()`, or `syms()`. Other functions are already loaded into dplyr.

Be careful when using `!!` without parentheses.